---
title: "Lab 5 Alissa Patterson"
author: "Alissa Patterson"
date: "10/26/2021"
output: 
  html_document: 
    toc: yes
    theme: cosmo
    number_sections: yes
    code_folding: hide 
---

```{r setup, include=TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

stage and commit. type commit message you want to be associated with changes. Press commit. Pull (down arrow). Then Push. Then when you go to github repo and refresh. 

calculate confidence interval for single sample

# Confidence intervals 

In this section we'll learn to use the 't.test)' function as a way to return the condfidence interval using the t-distribution.

## Example 1 

Create a mock sample of observation values, and find the confidence interval. Then we'll learn how to report that. 

```{r}
pika_mass <- c(82, 94, 110, 5000, 86, 99, 102) 

pika_mean <- mean(pika_mass)

pika_sd <- sd(pika_mass)

pika_ci <- t.test(pika_mass)
```
degrees of freedom is measure of sample size or number of variables 
## Risky and Tedious way:

Mean pika mass at Paiute Pass is 91.9 g $\pm$ 13.5 g (mean $\pm$ 1 standard deviation, n = 7) with a 95% confidence interval of [79.4, 104.4] g.

don't manually type or copy and paste values

command return to run single line

In-line code referencing to stored objects!

Use a single backtick on either side of a lowercase r to create an in-line bit of code

## A high initial investment, but safer and better way

Mean pika mass at Paiute Passs is 
'r round(pika_mean, 1)' g $\pm$ 'r round(pika_sd, 1)' g (mean $\pm$ 1 standard deviation, n = 'r
Length(pika_mass)') with a 95% confidence interval of ['r pika_ci$conf.int[1]', 'r pika_ci$conf.int]


# Two-sample t-test

Use a two-sample t-test to test the null hypothesis that samples were drawn from populations with the same mean (difference in means = 0).

- H0: Means difference = 0
- HA: Means difference is NOT - 0

this is a two-sided t-test beacuse no directionality is implied.

## Example 1 

```{r}
oak_trees <- c (29, 19, 22, 30, 35, 16, 20, 7)

pine_trees <- c(48, 32, 41, 40, 56, 70, 30, 19, 20)
```

Is there a significant difference in mean heights for oak and pine trees? 

```{r}
trees_t <- t.test(oak_trees, pine_trees) 
```

The p-value of 'r trees_t$p.value' means that...

1.8% chance of finding sample means *at least as different as those I found* by random chance if they were drawn from populations with the same mean height. 

## Example 2

Using the 'mpg' dataset to compare city gas mileage for SUVs and compact cars.

make data set that only contains obvs where class is suv or compact. filter keeps rows.

```{r}
compact_suv <- mpg %>%
  filter(class %in% c("suv", "compact"))
```


```{r}
ggplot(data = compact_suv, aes(x = cty)) + 
  geom_histogram(bins = 12)
facet_wrap(~ class)

ggplot(data = compact_suv, aes(sample = cty)) + 
  geom_qq() + 
  facet_wrap(~ class)
```

```{r}
car_stats <- compact_suv %>%
  group_by(class) %>%
  summarize(mean_city = mean(cty),
sd_city = sd(cty),
sample_size = n())
```

```{r}
compact_sample <- compact_suv %>%
  filter(class == "compact") %>%
  pull(cty)

cars_t <- t.test(compact_sample, suv_sample)

```


Mean gas mileage for compact cars and SUVs differs significatnly (t('r' cars_t$parameter') = 'r cars_t$statistic', p < 0.001, $\alpha$ = 0.05)
